---
title: "DVN Assignment 1 - Ggplot2"
output: html_notebook
---

Assignment 1 for Data Visualisation and Narratives requires the use of three
different visualisation tools over three weeks, presumably as part of a journey
of discovery through different visualisation tools. For this first tool I'm 
going to use something I already know reasonably well - Hadley Wickham's ggplot2
package for R.

The dataset I will be using for this analysis is something I've been thinking of
looking at for a while: the Australian Road Deaths Database. I've heard a lot of
statistics in the media which are presumably derived from this dataset, and I've
long suspected that there are some inaccurate claims floating around the place,
so I thought it was worth taking a look.

First things first, I need to grab the dataset.

```{r, warning=FALSE, message=FALSE}
library(httr)
library(readr)
request <- GET("https://bitre.gov.au/statistics/safety/files/Fatalities_Jan2017.csv")
text <- content(request, as="text")
raw_data <- read_csv(text)
```

Then I need to do some tidying up.

```{r}
library(dplyr)
library(lubridate)
fatal <- 
  raw_data %>% 
  mutate(Date = make_datetime(Year, Month, Day, Hour, Minute),
         Bus_Involvement               = Bus_Involvement == "Yes",
         Heavy_Rigid_Truck_Involvement = Heavy_Rigid_Truck_Involvement == "Yes",
         Articulated_Truck_Involvement = Articulated_Truck_Involvement == "Yes",
         State      = as.factor(State),
         Crash_Type = as.factor(Crash_Type),
         Road_User  = as.factor(Road_User),
         Gender     = as.factor(Gender)) %>% 
  select(-Day, -Month, -Year, -Hour, -Minute)
```

The first visualisation I will use is barely even a visualisation, but I would 
argue that a table of summary statistics counts! It certainly helps to guide the
narrative by drawing the eye to interesting patterns.

```{r}
fatal %>% 
  select(-CrashID, -Date) %>% 
  summary()
```
